A multi-species benchmark for training and validating large scale mass spectrometry proteomics machine learning models

### 2024-05-15 doc/nine-species/nine-species.tex
Add descriptions of studies

\fixme{Describe each study}

We created a new version of the nine-species benchmark originally described by Tran {\em et al.} \cite{tran2017denovo}
To do so, we downloaded the RAW files from the same nine PRIDE projects (Supplementary Table~\ref{tab:benchmark}) and converted them to MGF format using the ThermoRawFileParser v1.3.4.
We also downloaded the corresponding nine Uniprot reference proteomes and constructed a Tide index for each one, using Crux version 4.1.
For one species (\textit{Vigna mungo}), no reference proteome is available, so we used the proteome of the closely related species \textit{Vigna radiata}.

@@

\fixme{
\begin{itemize}
\item Annotated MGF files for the nine-species benchmark. \fixme{Do we provide them separately by run, or just record in the MGF records what run each one came from?}
\item Intermediate files for nine-species (genome FASTA, tide-search results, Percolator files).
\end{itemize}
}

>>>>

We created a new version of the nine-species benchmark originally described by Tran {\em et al.} \cite{tran2017denovo}

\begin{enumerate}
\item Paiva \textit{et al.}\ investigated the protein expression response of the cowpea plant (\textit{Vigna unguiculata}) to infection by \textit{Cowpsea severe mosaic virus} (CSMV) by carrying out label-free proteomic analysis of cowpea leaves that were inoculated with CSMV compared to mock inoculation controls \cite{paiva2016label}.
\item Nevo \textit{et al.}\ studied a rare autosomal recessive lysosomal storage disorder, cystinosis, by carrying out SILAC proteomic analysis of engineered mouse cell lines that harbor a known pathenogenic mutation of the causative gene, \textit{CTNS} \cite{nevo2017impact}.
\item Cassidy \textit{et al.}\ evaluated two different analytical approaches for carrying out full proteome analysis while identifying short open reading frames: high/low pH reversed phase LC-MS bottom-up approach and a semi-top-down strategy involving separation of proteins a GelFree system followed by digestion and LC-MS analysis \cite{cassidy2016combination}. The experiments were carried out using the methane producing archaeon \textit{Methanosarcina mazei}.
\item Reuss \textit{et al.}\ carried out proteomic analyses on a series of minimized strains of the model bacterium, \textit{Bacillus subtilis}, with genomes reduced by $\sim$36\% \cite{reuss2017large}.
\item Petersen \textit{et al.}\ performed proteomic analysis of \textit{Candidatus endoloripes}, which are bacterial symbionts of the \textit{Lucinidae} family of marine bivalves \cite{petersen2016chemosynthetic}.
\item Mata \textit{et al.}\ characterized the proteome of the tomato pericarp at its ripe red stage \cite{mata2017depth}.
\item Seidel \textit{et al.}\ analyzed the global proteomic stress response in wildtype and two yeast knockout strains for the gene PBP1  \cite{seidel2017quantitative}.
\item Hu \textit{et al.}\ studied honeybees that exhibit a suite of behaviors (\textit{Varroa} sensitive hygiene---VSH) associated with infection with the \textit{Varroa destructor} virus \cite{hu2016proteome}.  Proteomic analysis was carried out on mushroom bodies and antennae of adult honeybees with and without VSH.
\item Cypryck \textit{et al.} characterized extracellular vesicles released from human primary macrophages after infection with influenza A viruses \cite{cypryk2017proteomic}.
\end{enumerate}
All nine studies were performed using a Thermo Scientific Q Exactive mass spectrometer.

% The PXD entry says the species is "Vigna mungo (Rice bean) (Black gram); NCBI TaxID: 3915;" but the paper talks about Vinga unguiculata.

To do so, we downloaded the RAW files from the same nine PRIDE projects (Supplementary Table~\ref{tab:benchmark}) and converted them to MGF format using the ThermoRawFileParser v1.3.4.
We also downloaded the corresponding nine Uniprot reference proteomes and constructed a Tide index for each one, using Crux version 4.1.
For one species (\textit{Vigna mungo}), no reference proteome is available, so we used the proteome of the closely related species \textit{Vigna radiata}.

@@

The dataset contains files resulting from various stages of the benchmark:
\begin{itemize}
\item Spectrum files in MGF format, produced by msconvert.
\item Reference proteome files in FASTA format, downloaded from Uniprot.
\item Search results files for both targets and decoys, in tab-delimited format, produced by Tide.
\item PSM-level Percolator results files for targets, in tab-delimited format.
\item Annotated MGF files for all three versions of the benchmark (overlap, main, and balanced).
\end{itemize}
Also included are log files for the steps of the analysis pipeline carried out using Crux \cite{park2008rapid} (Tide indexing, Tide search, and Percolator).
